%\newpage
%\setcounter{page}{1}
%\setcounter{section}{0}

\input{header}


\begin{document}



\begin{center}
{\LARGE Crypto Lab -- One-Way Hash Function and MAC}
\end{center}

\copyrightnotice

\newcounter{task}
\setcounter{task}{1}
\newcommand{\tasks} {\bf {\noindent (\arabic{task})} \addtocounter{task}{1} \,}

\section{Overview}

The learning objective of this lab is for students to get familiar with
one-way hash functions and Message Authentication Code (MAC). After 
finishing the lab, in addition to gaining a deeper undertanding
of the concepts, students should be able to 
use tools and 
write programs to generate one-way hash value and MAC for 
a given message.


\section{Lab Environment}
The lab is started from the Labtainer working
directory on your Docker-enabled host, e.g., a Linux VM.
From there, issue the command:
\begin{verbatim}
    labtainer onewayhash
\end{verbatim}
The resulting virtual terminals will include a display of
a bash shell.  The openssl package
and other software described below are pre-installed
on the system.


\section{Lab Tasks}

\subsection{Task 1: Generating Message Digest and MAC}

In this task, we will play with various one-way hash algorithms. 
You can use the following {\tt openssl dgst} 
command to generate the hash value for a file. To see the manpages, you can 
type {\tt man openssl} and {\tt man dgst}.

\begin{Verbatim}[frame=single] 
  % openssl dgst dgsttype filename
\end{Verbatim} 

Please replace the {\tt dgsttype} with a specific one-way hash algorithm,
such as {\tt -md5}, {\tt -sha1}, {\tt -sha256}, etc.
And replace {\tt filename} with {\tt filetodigest.txt}, which is in your home directory.
In this task, you should try at least 3 different algorithms,
and describe your observations. 
You can find the supported one-way hash algorithms 
by typing {\tt "openssl dgst -h"}  NOTE: the list of algorithms included in the manpages is not correct.


\subsection{Task 2: Keyed Hash and HMAC}

In this task, we would like to generate a keyed hash (i.e. MAC)
for a file. We can use the {\tt -hmac} option (this option 
is currently undocumented,
but it is supported by {\tt openssl}). The following example
generates a keyed hash for a file using the HMAC-MD5 
algorithm. The string following the {\tt -hmac} option
is the key. 
\begin{Verbatim}[frame=single]
 % openssl dgst -md5 -hmac "abcdefg" filename 
\end{Verbatim} 

Please generate a keyed hash using HMAC-MD5, HMAC-SHA256, and
HMAC-SHA1 for any file that you choose. Please try several
keys with different length. Do we have to use a key 
with a fixed size in HMAC? If so, what is the key size? If 
not, why?


\subsection{Task 3: The Randomness of One-way Hash}


To understand the properties of one-way hash functions, we would like to
do the following exercise for MD5 and SHA256:

\begin{enumerate}
\item Create a text file named "edit-this-file.txt" of any length.
\item Generate the hash value  $H_1$ for this file using a specific hash algorithm.
\item Flip one bit of the input file.
      You can achieve this modification using {\tt hexedit}.
\item Generate the hash value $H_2$ for the modified file.
\item Please observe whether $H_1$ and $H_2$ are similar or not. Please describe
      your observations in the lab report. You can write a short program to count
      how many bits are the same between $H_1$ and $H_2$.

\end{enumerate}


\subsection{Task 4: One-Way Property versus Collision-Free Property}


In this task, we will investigate the difference between 
hash function's two perperties: one-way property versus collision-free property. 
We will use the brute-force method to see how long it takes to break
each of these properties. Instead of using {\tt openssl}'s command-line tools, 
you are required to write our own C programs to invoke the 
message digest functions in {\tt openssl}'s crypto library. 
A sample code can be found from 
\url{http://www.openssl.org/docs/crypto/EVP_DigestInit.html}. Please 
get familiar with this sample code.

Since most of the hash functions are quite strong against the brute-force
attack on those two properties, it will take us years to break them
using the brute-force method. To make the task feasible, we reduce 
the length of the hash value to 24 bits. We can use any one-way
hash function, but we only use the first 24 bits of the hash value
in this task. Namely, we are using a modified one-way hash function.
Please design an experiment to find out the following:

\begin{enumerate}
\item How many trials it will take you to break the one-way
property using the brute-force method? You should repeat your 
experiment for multiple times, and report your average number 
of trials.
\item How many trials it will take you to break the collision-free
property using the brute-force method? Similarly, you should 
report the average.
\item Based on your observation, which property is easier to 
break using the brute-force method?
\item (10 Bonus Points) Can you explain the difference in your 
observation mathematically?
\end{enumerate}

 

\section{Submission}
When the lab is completed, or you'd like to stop working for a while, run
\begin{verbatim}
    stoplab onewayhash
\end{verbatim}

from the host Labtainer working directory.  You can always restart the
Labtainer to continue your work.  When the Labtainer is stopped, a
zip file is created and copied to a location displayed by the stoplab
command.  When the lab is completed, send that zip file to the instructor.

You need to submit a detailed lab report to describe what you have
done and what you have observed; you also need to provide explanation
to the observations that are interesting or surprising. In your report,
you need to answer all the questions listed in this lab.




\end{document}


