Note on backward compatibility

A user may easily get a new container image that requires framework
script support not present in the user's environment, e.g., has
not yet done an update-labtainer.sh.  

A user may have an old container not supported by framework scripts
obtained via an update-labtainer.sh

\section{Lab Versions}
Substantive changes to an existing published lab should be made in
a new named lab.  A "substantive" change is defined as one that would
break any existing installation.  The sole exception is the case in 
which a newer lab image would require a Labtainers update.

Never change container names for existing labs.

Management of multiple versions of the same lab will be achieved entirely
by the user environment and what is distributed to it.

\subsection{Managing versions}
Labtainer updates will not remove old labs from installed systems, however 
they may stop distributing old labs.

If a user installation contains two versions of a lab, only the latest will
be displayed in the labtainer command -- unless the installation includes
containers for the older -- in which case both will be listed.

If a user starts an older version of a lab for the first time, warn the user
with a prompt.  Subsequently, note the existence of the newer lab without a prompt.

Lab names change with versions, e.g., centos-logv2.  (tab completion would really
help here).

Defining multiple labs as being different versions of the same lab is achieved
in a "version" file in the lab/config directory.  Absence of the file implies the
lab is its only version and is its base name. Otherwise, the file includes the 
lab basename and its version number, which is syntatically independent of the name
of the lab, e.g., lab names need not incorporate a "v2".  Version numbers are integers.
Lab basnames are ONLY used to distinguish versions of the same lab.   Students and
instructors name labs using their lab name, not their basename.

\subsection{Images dependent on framework versions}
The sole exception to the requirement for compatability of lab versions is the
case where we can request the user to update the labtainers framework in order
to run the new labs.  This concept is not limited to versions of the same lab,
it also pertains to changes to scripts in images that require corresponding
support from the framework.  An example might be a new argument to the
parameterize.sh script.  If new images expect this parameter, then the framework
must provide it, i.e., it must be updated.  (Backward compatability must still be
supported, i.e., old images MUST run with all new frameworks.)

These versions relationships will be managed crudely.  Once any new lab requires an
updated framework, all subsequent labs will require at least that framework version,
(whether it needs it or not).

These versions are managed by a version number imbedded in the build process.


\subsection{Base image replacements and changes}
Transitioning to new base images will force some users to wait for new large downloads
when starting a new lab.

Can we tell the student how large the download will be?  not likely.
But we can identify the lack of a base image on the student computer, and notify them
of that.

MOTIVE:  Reduce cumulative size of all base images needed to perform labs.
This means always rebuilding all labs whenever a base changes.

OPTIMIZE for size of the VM.

Roadmap:  
\begin{itemize}
\item Replace existing images with identical images having new version labels,
thus forcing an update of the framework.
\item New framework version will be sensitive to base image versions.

\item Re-tag all lab images to reflect an initial base image version.

\item Before a pull (or a update-labtainer.sh!), the framework will determine its 
base-image version, and pull the lab image versions consistent with that base image.

\item The tag associated with each lab image will reflect its base image
dependency.  The "latest" image version implies the latest base image version.

\item The initial base image version is 1.  All images will be tagged to reflect
that version.

\item Separate logic for ensuring student has necessary base for a lab, and prompt if 
a big download is needed.
\end{itemize}

NEVER PULL and old base.  NEVER build from an old base

If a base is updated, you MUST rebuild all labs.  



PROCEDURE
publish new framework
retag all images

rebuild base images and push

rebuild all images


HOW to run and test old images?

Keep an old VM!




